\chapter{Acknowledgements}
We would like to thank our supervisors from ITU, Gordon Calleja and Joel Anthony Lehman, for the help and guidance with the project and dissertation. 

\section{Team composition}
This project group consists of three members; Maxime Moze, Mats Stenhaug and Mikkel Stolborg. All three are, at the time of writing this paper, master students at the IT University of Copenhagen, class of 2016. Maxime Moze will be the game designer of the group, and his main roles will be to create the board game with the physical pieces and layout, come up with the rules, mechanics and flow of the game, along with prototypes and input to the programmers on how the connection between the app and board should be. He will also be the one in charge of giving feedback to the programmers as to what functionality should be in the app, and user friendliness of the app.
The programmers in this group are Mikkel Stolborg and Mats Stenhaug. They will be in charge of coming up with the procedural algorithms for generating the content along with programming the application itself, and making sure the app is capable of utilizing memory in order to generate more content to be presented to the players. Mainly the roles of the programmers will be to create the logic behind the procedural content generation, and utilizing the algorithms in ways that will benefit the game as a whole, and give the players the best possible experience of using the digital aspect of the game, i.e. the application.
Together we will be communicating back and forth and work close together to ensure that the development process will go as smoothly as possible, and that there will be as little communication errors as possible.