\section{Background}
\subsection{Basic definitions}
It is commonly admitted that there are as many definitions of games as there are players. People working in the game industry, researchers, and players still disagree on which key elements should a taxonomy of games be based on. The french sociologist Roger Caillois (1958, p.3) explains the difficulty of establishing an all-inclusive game classification:
\begin{quotation}
The multitude and infinite variety of games at first causes one to despair of discovering a principle of classification capable of subsuming them under a small number of well-defined categories. Games also possess so many different characteristics that many approaches are possible.
\end{quotation} 
With this statement, one could argue that defining or classifying game genres is not relevant, as there will always be examples that do not fit the definition, which will then need to be extended and will therefore move away from its original meaning. This paper will certainly not bring a clear and universal definition of the game genres relevant to its topic, as definitions and taxonomies are necessarily affected by their research purpose. However, basic definitions are needed to build an understandable framework and provide tools that will help when analysing the impact of the use of PCG in tabletop games, both on their design and creation and on the experience they enable.
Obviously, the implementation of digital components into tabletop games can already cause problems to some people, depending on their understanding on what tabletop games are. The following paragraphs will narrow down the definitions, in order to point out the key elements that will be referred to in the analysis of the contribution of PCG in Archipelago. Those key elements will be then used to point out existing games applicable to the topic of this paper.
\subsubsection{Tabletop Games}
The definition of "tabletop games" is the first one that must be refined in order to correctly comprehend the purpose of this paper. An example is needed as a reference in order to point out their key elements. Nat Levan, designer at Oakleaf Games gives his own definition of tabletop games. (Levan, 2014)
\begin{quotation}
The defining feature of a tabletop game is the need for components, not the need for a table. [...] so perhaps saying that a tabletop game is a game with a structured placement of components establishes a better requirement.
\end{quotation} 
"Tabletop games" is a generic term used to refer to a broad variety of game genres that are played on a table or the like. They range from card games, board games, not forgetting miniature games or dice games. The aim of defining "tabletop games" this way is also a way to exclude digital games. Although removing the need for a table from the definition seems rather counter-intuitive, designers make this assumption to include other simple games using components. Nat Levan uses the example of the game of "guess what I’m holding in my hand"(Levan, 2014). However, it seems that this approach is incomplete, as it overshadows the social aspect of tabletop games. It is understandable that tabletop games are not necessarily meant to gather people around a table to play. As an example, \textit{Magic: The Gathering} (MicroProse, 1997) is a digital version of the Wizards of the Coast’s trading card game of the same name. Moreover, websites allowing traditional tabletop games like \textit{Carcassone} (Klaus-Jürgen Wrede, 2000), \textit{Through the Ages: A Story of Civilization} (Vlaada Chvátil, 2006) or \textit{Race for the Galaxy} (Tom Lehmann, 2007) and others to be played online have made their appearance with the development of the internet. Some These games do not cease to be "tabletop games", even though the play experience is made different by changing the context in which they are played.\\
The next step of the process is to define what can be called "components", and why they have to be central in the game. Core components of tabletop games are generally thought as physical. They can be cards, tokens, pawns, or any object used as a central element in the game, which cannot be played without them.  As we know, tabletop games have generally evolved staying away from digital components. Exceptions exist though, and game makers have already experimented the use of technology to influence their design. The game \textit{Atmosfear} (Brett Clements, 1991) included a VHS tape showing "The Gatekeeper", who attributes rewards to the players, punishes them or gives them instructions at specific moments in the game, thus also acting as a timer. In that sense the VHS can be seen as a component in itself, like a deck of cards with a hourglass. 
Components could then be defined as interactive elements without which the game cannot be played. The central dimension of components is clearly understandable, but it is important to notice that their physical property has been excluded from the definition.\\
After this reflection, it is now possible to come up with a definition of what tabletop games are. Tabletop games cannot be narrowed to the simple use of physical components, neither must they only be played on a table; a more appropriate definition could be that tabletop games are games requiring the players to use central components in a space - real or virtual - representing a table or similar surface.
\subsubsection{Board Games}
"Board games" are generally considered as a subcategory of tabletop games. The Oxford Dictionaries (2016) provide the following definition:
\begin{quotation}
"A game that involves the movement of counters or other objects round a board."
\end{quotation}
While this is a rather broad definition, it can still be used to present key aspects of board games. The terms "Counters or other objects" refer to the core components that were previously mentioned as a key aspect of tabletop games. The new element refining the definition being the game board, which is the surface on which the players move pieces, place cards or tiles in order to achieve the goal of the game. As it has been explained beforehand, this surface is not necessarily physical and can be then represented on a digital device like a screen. In a board game, the board is of course a key component. It is the center of the attention of players, and can serve several purposes.\\
The board concentrates all the structural elements that the players should focus on in order to correctly play the game. It must support the play experience as much as put limitations on the way the game is played. It that sense it is as much an interface from which players get the more or less abstract information they need, as it is an obstacle forcing them to play according to the rules. In the game of \textit{Chess} for example, those properties are clearly identifiable, though the board is not the most detailed. It limits the movement of the pieces within a square of 8 units of length, and a trained eye can easily identify valuable information (i.e. movement possibilities or strategy planning) , thanks to the tiled structure of the board. One last property that can be identified, is that a board is rarely a core component in itself. During a game, the board only make sense when it is accompanied by other core components like tokens or cards. Boards have different structures and purposes, and listing their properties here would not be relevant. But from those identified earlier, it becomes possible to give a definition of "board games": "board games" are tabletop games using a game board and its properties as a core component.
\subsubsection{Collaborative Games}
Though their approach on the play experience is somewhat similar, researchers traditionally make the difference between cooperative and collaborative games. Nash (2002, cited in Zagal et al., 2006, p.25) argues that in cooperative games,the players have interests that are "neither completely opposed, nor completely coincident". In contrast,"in a collaborative game, all the participants work together as a team, sharing the pay-offs and outcomes; if a player wins or loses, everyone wins or loses"(Zagal and Rick,2006, p.25). This differentiation between "cooperation" and "collaboration" makes a big difference in the play experience. As an example, \textit{Diplomacy} (Allan B. Calhamer, 1959) encourages players to create alliances between them, and betray them when it is needed. However, in \textit{Forbidden desert} (Matt Leacock 2013), players have to manage the resources they individually possess, and put their skills at the service of the team. 
The differentiation is usually made in the game's winning condition. Cooperative games are still based on competition, whereas collaborative games have only winners or only losers. Collaboration involves more open discussions among the players and exclude the necessity to voluntarily hide information. 
\subsection{Defining "Hybrid Board Games"}
"Hybrid games" is an expression that has many different definitions. Therefore, there is not a universal understanding of what hybrid games are. Some people simply call "hybrid games" games that are a combination, or hybridization, of two or more different genres of games. The interest of that definition is that it removes the analogue or digital aspect of games and simply focuses on the experience they provide. It needs to be refined though, because many games can be seen as a combination of genres, as game designers often build upon already existing game mechanics, trying to adapt them in another context. In this paper, "hybrid board games" will be referred as board games (and all their properties previously defined) using one or several digital components. One criteria needs to be added though: as it has been already mentioned, the digital or analogue nature of the components or the board should not be what defines a hybrid game. A game is much more than its components, gameplay mechanics or even its rules. A game can be defined by the experience that the combinations of all these factors enable. Therefore, only board games including digital components in order to enable an experience that cannot exist without using the computing power of digital devices will be considered as "hybrid board games" in this paper. Games that are only using digital devices to simulate analogue game components (deck of cards or board for example) without providing any new feature are not considered hybrid games. Several games fitting this definition have already been created. The three examples presented later show the potential of this approach of hybrid games, each in its own way.
\subsubsection{Hybrid Experiments: False Prophets (Mandrik and Maranan, 2000)and STARS (Magenkurth Stenzel and Prante, 2003)}
People have also started to refer to "hybrid games" as games that try to close the gap between analogue and digital games. False Prophets is an experimental game that explores "the space between board games and video games, leveraging (sic)the advantages of both"(2000, Mandrik and Maranan). This experiment was a technical challenge more than a game design experiment. Mandryk and Maranan created a tabletop display using sensors in order to recreate an interface. A map is projected on a surface containing infra-red sensors, which can detect the position of the tokens in the displayed tiles. Originally, the map is hidden. It is only when the game's computer detects players approaching its borders that the map expands. The interest of this experiment is that it tries to exploit the assets of both game boards and computers. It exploit the setting of having an interface around which players can gather, and computation can be used to do complex calculation that would be very complex to do for humans. In this case, the manifestation of that is the map that gets revealed dynamically. While fully analogue games  that possess this feature, like \textit{Zombies!!!}(2001, Todd Breitenstein and Kerry Breitenstein) already exist, the sensor interface could allow many possibilities for game mechanics. The creators of False Prophets mentioned having "playing pieces that store data, interpret contextual information, and travel continually with the players"(Mandrik and Maranan, 2000).
\\\\
STARS is another example of the integration of digital devices in the traditional way of playing tabletop games. However, this project is an attempt to create an "augmented tabletop game platform", which could be usable with multiple games. The platform features an interactive touch-sensitive screen, which displays the content of the game board as well as it manages the interactive playing pieces. A camera is situated over the table and tracks the position of other playing pieces. STARS also features a wall display and other digital components like a small Personal Digital Assistant (used by the players to secretly communicate and store personal information related to the game), and audio devices. The experiment is an attempt to exploit a ubiquitous computing environment (an environment exploiting multiple devices) to allow the creation of more complex game rules, and recreate a game world with more fidelity without slowing down the game flow with multiple complex calculation. It also allows multiple game sessions, as well as game persistence. Although this experiment looks similar to False Prophets, its interest is that it exploits multiple devices and their affordances in order to create a platform which can then be used for multiple games. It opens to the possibility of creating other tabletop games using digital devices as core components and enable new experiences. One could argue that the quantity of available information (on the board, the wall display and the personal digital assistant) might get in the way of play by making it too complex. But this invention is interesting in its approach of hybrid board games: it uses the potential of digital devices as core components, enabling new play experiences that would hardly be possible otherwise.
\subsection{Procedural Content Generation}
"Procedural content generation (PCG) refers to the algorithmic creation of content. It allows content to be generated automatically, and can therefore greatly reduce the increasing workload of artists." (Rolan van der Linden, et al., Procedural generation of dungeons, March 2014).
This quote tells us that PCG does not really have anything to do with the playing of the game it self, rather the contents within it, hence the name procedural "content" generation. It is becoming more and more accepted within larger companies, as a way to lighten workload of other artists, and its uses range between various fields.
To build on what R. van der Linden, et al. say in their paper, about reducing workloads, PCG is a great tool to create templates for contents, based on data from designer, artists, etc. A technical term within the PCG world is Computer Assisted Design. This means that the computer does the generation initially, produces a result, and that result is then used by the other artists as a template to continue building from. So if the need to generate a map, PCG can do that by taking in certain parameters into account, and then it is up to the "humans" to look over and see whether or not it is pleasing enough, fun enough, or in general just acceptable. It is in the minds of humans that creativity lies. \\

Other aspects within the PCG world, which is becoming more and more accepted and used, is "Experience Driven" PCG. This concept is taking into account the previous events and data gathered within the game or application. By using gathered data, the newly generated contents can be altered to better fit with the way the previous content was used. Say you have a map in a first person shooter (FPS), and the initial map was generated statically by PCG, and not by E.D.PCG. The first map could be a very good one, by all means, but the program has no way of actually knowing how it will be used. Then by gathering data from an actual live run of the program, the PCG algorithm can then be altered to take into account frequent patterns that may occur, what open areas are most used, places with high chance of death, etc. Basically by combining some form of data mining with PCG and have a clever way of integrating them based on experience, you get a smarter way of generating content, and the content will have more "meaning" as it is in fact based on real life events and occurrences.\\

Genetic Algorithms
PCG is a very broad term, and it is becoming even broader. In the field of computer science, combining different areas into one program is not uncommon. This goes for PCG as well. As described above, PCG can be combined with data mining in order to create experience driven PCG. But there are many more things you can do, for instance, taking use of artificial intelligence. Specifically genetic algorithms. A genetic algorithm has the main function of taking a given part, called a Genome, and mutating parts of if, the chromosomes, and returning the changed values. However in order to select which mutated genome should be used, a fitness function is needed. What these functions do, is to test the generated genomes against parts of the program that is applicable. By scoring the genomes performance on the parts tested up against, you are able to select the highest scoring one. The best one. 

Sometimes, however, the highest scoring genome is not necessarily the best one. Sometimes combining a lower scoring genome with a higher scoring one can diversify the results, and the outcomes can be completely different, educational, and sometimes even better than before. Having this in mind when using this algorithm, and also when generally just combining different methods together, can be a smart move. Sometimes having a novelty search will grant the results that would not have been possible if only using a best-fit selection.

\subsubsection{Procedural Content Generation in Game Design}

PCG being a programming concept allowing the rapid creation of content, it has been mostly used for video games. There are counter examples to that statement though. In films for example, PCG can be used to rapidly generate complex scenes. The battle scenes \textit{The Lord Of The Rings: The Two Towers} (2002) and its sequel \textit{The Lord Of The Rings: The Return of the King} (2003) for example have been made possible thanks to an agent generator software called \textit{Massive} (2001, Weta Digital), which procedurally generates intelligent and independent agents in order to simulate large-size battles. This paper will leave that case aside though and focus in the next paragraphs on PCG-based games in order to point out the Game Design concepts that are enhanced thanks to the use of PCG. The basic concepts have been identified by researchers (2011 Smith et al.)\\
The most obvious aspects reinforced by PCG that are pointed out in this article are replayability and adaptability, which are considered interconnected.\\
Replayability is a concept used to describe the fact that a player can still discover new aspects of a game even after finishing it the first time. The paper refers to the game \textit{Rogue}(Toy, Wichman and Arnold, 1980), as one of the earliest and most famous example of PCG-based games. The basic concept of \textit{Rogue} is simple and has given birth to a new game genre - also giving it its name: Rogue-like games. The principle of Rogue-like games is the following: the player explores his environment (a dungeon in the case of \textit{Rogue}), collects items (like equipment) to become stronger and defeat more and more dangerous enemies. In Rogue-like games, the environment (and thus the collectible items and the enemies) are procedurally generated which ensures that the players are really unlikely to face the same situation when going through the game. This shows how PCG can enhance the replayability of a game.\\ 
In the same study, adaptability is referred as "the use of procedural content generation to adjust content in reaction to player actions or skill levels". An example can be found in the video game \textit{FTL: Faster Than Light} (Subset Games, 2011). This game - which uses Rogue-like game mechanics - is based on procedurally generated events that players   have to overcome, sometimes by choosing between two different solutions. The selected option will then influence the next events generated by the game. In that case, the player's actions are the parameters used by the algorithm to generate the content. In other words the game content adapts itself to the player's action.\\
A distinction is also made between "PCG systems augment traditional mechanics and those which enable new mechanics entirely" (2011, Smith et al.). That distinction is made to differentiate the impact of PCG on the game's mechanics and the effect on the play experience. The example of \textit{Borderlands} (Gearbox Software 2009) is given in the paper to demonstrate that PCG does not change the core mechanics of the game itself - a first-person shooter game in which the players collect more and more powerful procedurally generated weapons. However, the play experience is vastly affected by the perceived randomness that PCG can enable. The excitement when waiting for the containers holding weapons, or the sudden burst of adrenaline when a powerful weapon appears could not be reproduced without the use of PCG. The Game Designers seem to have also enhanced that aspect in the multi-player mode by making the generated items collectible by any player, even players that have never taken part in the battle (this mechanic is generally called "ninja-loot"). Another example that comes to mind is the game \textit{Diablo} (Blizzard North 1996). It uses the same approach to procedurally generate equipment, but the emphasis is put on the optimization of the character's attributes and skills. This use of PCG is opposed to game mechanics based on PCG: \textit{Galactic Arms Race} (Evolutionary games 2014) uses a mechanic in which the more a weapon is used, the more the chances of getting a weapon having similar properties increase. This mechanic is used to base the gameplay on the player's choices, as it becomes important to use the weapons with the preferred attributes, in order to get similar weapons later in the game. In that sense, the mechanic can be considered PCG-based.\\
Finally, this mechanic also demonstrates the last aspect of PCG-based Game Design described in the article, which is the player's control over the game's content. This means that the player has an influence over the PCG system in the game. In theory, the players input are often used by the algorithm to use the new content, so the influence of the player is real. However, in order to have a certain control over the game's content, a player must be aware that she influences the algorithm with her choices.

